For Finitary Induction-Induction, \\ Induction is Enough

### 2020-05-02 paper/paper.tex
sec2

\begin{definition}[Signature algebra, $\SignAlg$]\label{def:algebra}
  In a model of ETT, a signature algebra is an iterated $\Sigma$ type
  consisting of the following four (families of) sets, 17 operations
  and 18 equalities.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\
  & \Con && : \Set \\
  & \Ty  && : \Con\ra\Set \\
  & \Sub  && : \Con\ra\Con\ra\Set \\
  & \Tm  && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \\
  & \id && : \Sub\,\Gamma\,\Gamma \\
  & \blank\circ\blank && : \Sub\,\Theta\,\Delta\ra\Sub\,\Gamma\,\Theta\ra\Sub\,\Gamma\,\Delta \\
  & \ass && : (\sigma \circ \delta) \circ \nu = \sigma \circ (\delta \circ \nu) \\
  & \idl && : \id\circ\sigma = \sigma \\
  & \idr && : \sigma\circ\id = \sigma \\
  & \blank[\blank] && : \Ty\,\Delta\ra\Sub\,\Gamma\,\Delta\ra\Ty\,\Gamma \\
  & \blank[\blank] && : \Tm\,\Delta\,A\ra(\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma]) \\
  & [\id] && : A[\id] = A \\
  & [\circ] && : A[\sigma \circ \delta] = A[\sigma][\delta] \\
  & [\id] && : t[\id] = t \\
  & [\circ] && : t[\sigma \circ \delta] = t[\sigma][\delta] \\
  & \boldsymbol{\cdot} && : \Con \\
  & \epsilon && : \Sub\,\Gamma\,\boldsymbol{\cdot} \\
  & {\boldsymbol{\cdot}\eta} && : (\sigma : \Sub\,\Gamma\,\boldsymbol{\cdot}) \ra \sigma = \epsilon \\
  & \blank\ext\blank && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \blank,\blank && : (\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma])\ra\Sub\,\Gamma\,(\Delta\ext A) \\
  & \pi_1 && : \Sub\,\Gamma\,(\Delta\ext A)\ra \Sub\,\Gamma\,\Delta \\
  & \pi_2 && : (\sigma : \Sub\,\Gamma\,(\Delta\ext A))\ra \Tm\,\Gamma\,(A[\pi_1 \sigma]) \\
  & {\pi_1\beta} && : \pi_1(\sigma, t) = \sigma \\
  & {\pi_2\beta} && : \pi_2(\sigma, t) = t \\
  & {\pi\eta} && : (\pi_1\,\sigma, \pi_2\,\sigma) = \sigma \\
  & {,\circ} && : (\sigma, t)\circ\delta = (\sigma\circ\delta, t[\delta]) \\
  & \rlap{(2) Universe} \\
  & \U && : \Ty\,\Gamma \\
  & \El && : \Tm\,\Gamma\,\U\ra\Ty\,\Gamma \\
  & {\U[]} && : \U[\sigma] = \U \\
  & {\El[]} && : (\El\,a)[\sigma] = \El\,(a[\sigma]) \\
  & \rlap{(3) Inductive parameters} \\
  & \Pi && : (a:\Tm\,\Gamma\,\U)\ra\Ty\,(\Gamma\ext\El\,a)\ra\Ty\,\Gamma \\
  & \blank\oldapp\blank && : \Tm\,\Gamma\,(\Pi\,a\,B)\ra (u : \Tm\,\Gamma\,(\El\,a))
  \ra \Tm\,\Gamma\,(\El\,(B[\id,\,u])) \\
  & {\Pi[]} && : (\Pi\,a\,B)[\sigma] = \Pi\,(a[\sigma])\,(B[\sigma\circ\p,\q]) \\
  & {\oldapp[]} && : (t\oldapp\alpha)[\sigma] = (t[\sigma])\mathop{\oldapp}(\alpha[\sigma]) \\
  & \rlap{(4) External parameters} \\
  & \Pim && : (T:\Set)\ra(T\ra\Ty\,\Gamma)\ra\Ty\,\Gamma \\
  & \blank\appm \blank && : \Tm\,\Gamma\,(\Pim\,T\,B)\ra(\alpha:T) \ra\Tm\,\Gamma\,(B\,\alpha) \\
  & {\Pim[]} && : (\Pim\,T\,B)[\sigma] = \Pim\,T\,(\lambda \alpha.(B\,\alpha)[\sigma]) \\
  & {\appm[]} && : (t\appm\alpha)[\sigma] = (t[\sigma])\mathop{\appm}\alpha
\end{alignat*}
Given an $M:\SignAlg$, we denote its components as $\Con^M$, $\Ty^M$,
$\Sub^M$, $\Tm^M$, $\id^M$, and so on.
\end{definition}

\begin{definition}[Abbreviations]\label{def:abbrevs}
  For any signature algebra, we use
  $\wk : \Sub\,(\Gamma\ext A)\,\Gamma$ to mean $\pi_1\,\id$. We
  recover de Bruijn indices by setting $0:= \pi_2 \,\id$ and
  $1+n := n[\wk]$. $\Pi\,a\,(B[\wk])$ is abbreviated by $a\Ra B$,
  $\Pim\,T\,(\lambda\_.B)$ by $T\Ram B$.
\end{definition}

The definition of morphism is determined by the notion of algebra, it
can be computed following \cite{large_inf_qiit}, but we include it
here for completeness.
\begin{definition}[Signature morphism, $\SignMor$]\label{def:morphism}
  A morphism from signature algebras $M$ to $N$ denoted
  $\SignMor\,M\,N$ consists of four functions and 17 equalities
  expressing that the functions preserve the operations of the two
  algebras. We use the same naming as in \Cref{def:algebra} and use
  superscripts to denote which algebra is meant.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\
  & \Con && : \Con^M && \ra\Con^N \\
  & \Ty  && : \Ty^M\,\Gamma && \ra\Ty^N\,(\Con\,\Gamma) \\
  & \Sub  && : \Sub^M\,\Gamma\,\Delta && \ra\Sub^N(\Con\,\Gamma)\,(\Con\,\Delta) \\
  & \Tm  && : \Tm^M\,\Gamma\,A && \ra\Tm^N\,(\Con\,\Gamma)\,(\Ty\,A) \\
  & \id && : \Sub\,\id^M && = \id^N \\
  & {\circ} && : \sigma\circ^M\delta && = \Sub\,\sigma\circ^N\Sub\,\delta \\
  & [] && : A[\sigma]^M && = \Ty\,A[\Sub\,\sigma]^N \\
  & [] && : t[\sigma]^M && = \Tm\,t[\Sub\,\sigma]^N \\
  & \boldsymbol{\cdot} && : \Con\,\boldsymbol{\cdot}^M && = \boldsymbol{\cdot}^N \\
  & \epsilon && : \Sub\,\epsilon^M && = \epsilon^N \\
  & \ext && : \Con\,(\Gamma\ext^M A) && = \Con\,\Gamma\ext^N\Ty\,A \\
  & , && : \Sub\,(\sigma,^M t) && = \Sub\,\sigma ,^N\Tm\,t \\
  & \pi_1 && : \Sub\,(\pi_1^M\,\sigma) && = \pi_1^N\,(\Sub\,\sigma) \\
  & \pi_2 && : \Tm\,(\pi_2^M\,\sigma) && = \pi_2^N\,(\Sub\,\sigma) \\
  & \rlap{(2) Universe} \\
  & \U && : \Ty\,\U^M && = \U^N \\
  & \El && : \Ty\,(\El^M\,a) && = \El^N\,(\Tm\,a) \\
  & \rlap{(3) Inductive parameters} \\
  & \Pi && : \Ty\,(\Pi^M\,a\,B) && = \Pi^N\,(\Tm\,a)\,(\Ty\,B) \\
  & \oldapp && : \Tm\,(t\mathbin{{\oldapp}^M} u) && = \Tm\,t\mathbin{{\oldapp}^N}\,\Tm\,u \\
  & \rlap{(4) External parameters} \\
  & \Pim && : \Ty\,(\Pim^M\,T\,B) && = \Pim^N\,T\,(\lambda\alpha.\Ty\,(B\,\alpha)) \\
  & \appm && : \Tm\,(t\mathbin{{\appm}^M}\alpha) && = \Tm\,t\mathbin{{\appm}^N}\alpha
\end{alignat*}
Given an $f:\SignMor\,M\,N$, we denote its first four components just
by $f_\Con$, $f_\Ty$, $f_\Sub$, $f_\Tm$ or just write $f$ if it is
clear which one is meant.
\end{definition}

\begin{definition}[IIT signature]
  An IIT signature is a context in an arbitrary signature algebra
  which is compatible with morphisms:
  \begin{alignat*}{6}
    & \Sign :=\,\, && \big({sig}:(M:\SignAlg)\ra\Con^M\big)\times \\
    & && \big((M\,N:\SignAlg)(f:\SignMor\,M\,N)\ra f_\Con\,({sig}\,M) = {sig}\,N \big).
  \end{alignat*}
\end{definition}
The first component is the Church encoding of $\Con$, the second is
the compatibility condition introduced by Awodey, Frey and Speight
\cite{DBLP:conf/lics/AwodeyFS18}. We don't need impredicativity, as we
don't require that $\Sign$ is the initial signature algebra. The
compatibility condition says that if we obtain an $M$-context using
${sig}$ in model $M$ and then we transport it to $N$ using $f$, we get
the same $N$-context as directly applying ${sig}$ to $N$.

@@

\begin{example}[Example signatures]\label{ex:signatures}
  Using informal notation with variable names, the signature for
  natural numbers would be
  \[
    \boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N.
  \]
  This signature is a context of length three. We start with the empty
  context $\boldsymbol{\cdot}$, then we declare a sort $N:\U$, then we
  declare an operator producing an element of $N$ denoted by $\El\,N$,
  and finally we declare an operator which takes as input an element
  of the sort $N$ and produces an element of the same sort. Note the
  assymetry of the function type $\Ra$: on the left there needs to be
  an element of $\U$, while on the right any type can appear
  (including another function type). This ensures strict positivity of
  the operators.

  Our theory of signatures uses de Bruijn indices instead of variable
  names, so we rewrite the above signature to
  \[
    \boldsymbol{\cdot}\,\ext\,\U\,\ext\,z : \El\,0\,\ext\,s : 1 \Ra \El\,1,
  \]
  and as we $\Sign$ is defined formally as a pair of a Church encoding
  together with a compatibility condition, our final definition for
  the signature for natural numbers is the following. The
  compatibility condition is written using equational reasoning.
  \begin{alignat*}{5}
    & \rlap{$(nat,natc) := $} \\
    & \hspace{1em} && \big( && \rlap{$\lambda M.(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M),$} \\
    & && && \lambda M\,N\,f\,.\, && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M) = \\
    & && && && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M)\ext^N f_\Ty\,(1^N\Ra^N \El^N\,1^N) = \\
    & && && && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M)\,\ext^N\,f_\Ty\,(\El^M\,0^M)\ext^N f_\Tm\,1^N\Ra^M f_\Ty\,(\El^N\,1^N) = \\
    & && && && f_\Con\,\boldsymbol{\cdot}^M\,\ext^N\,f_\Ty\,\U^M\,\ext^N\,\El^N\,(f_\Tm\,0^M)\ext^N 1^M\Ra^M \El^M\,(f_\Tm\,1^N) = \\
    & && && && \boldsymbol{\cdot}^N\,\ext^N\,\U^N\,\ext^N\,\El^N\,0^N\,\ext^N\,1^N \Ra^N \El^N\,1^N \big)
  \end{alignat*}
  The first component builds the context describing natural numbers in
  $M$, the second one uses the fact that $f$ is a morphism, that is,
  it preserves all operations.

  Lists with elements of a given $T : \Set$ type are given by the
  following signature. Here we use the function space with external
  domain $\Ram$ to include a $T$ in the signature. For readability, we
  omit the $\lambda$ and the superscripts and we don't write the
  compatibility condition. On the right we list the same signature
  with variable names.
  \begin{alignat*}{10}
    & \boldsymbol{\cdot}\,\ext\,\U\,\ext\,\El\,0\,\ext\,T\Ram 1\Ra \El\,1 \hspace{3.56em} \boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,cons : T\Ram L\Ra \El\,L
  \end{alignat*}

  The $\Con$--$\Ty$ example from \Cref{sec:intro} is specified as
  follows.
  \begin{alignat*}{5}
    & \boldsymbol{\cdot} \,\ext                                                 && \boldsymbol{\cdot} \,\ext                                                                \\
    & \U\,\ext                                                                  && Con && : \U\,\ext                                                                        \\
    & 0\Ra\U\,\ext                                                              && Ty && : Con\Ra\U\,\ext                                                                   \\
    & \El\,1\,\ext                                                              && empty && : \El\,Con\,\ext                                                                \\
    & \Pi\,2\,(2\oldapp 0 \Ra \El\,3)\ext                                       && ext && : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \Ra \El\,Con)\ext                         \\
    & \Pi\,3\,(\El\,(3\oldapp 0))\,\ext                                         && U && : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))\,\ext                                 \\
    & \Pi\,4\,(\El\,(4\oldapp(2\oldapp 0\oldapp (1\oldapp 0)))) \hspace{4.79em} && El && : \Pi\,(\Gamma : Con)\,(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp (U\oldapp\Gamma))))
  \end{alignat*}
\end{example}

@@

\begin{definition}[The set signature algebra $\A$]\label{def:a}
  $\A : \SignAlg$ is given by the first $^\A$ components of $\ADS$
  (\Cref{def:ads}), that is, $\Con^\A := \Set$,
  $\Ty^\A\,\Gamma := \Gamma\ra\Set$,
  $\Sub^\A\,\Gamma\,\Delta := \Gamma\ra\Delta$, and so on. Note that
  there is a morphism from $\ADS$ to $\A$ defined by $\blank^\A$ at
  each component, which we also denote by
  $\blank^\A : \SignMor\,\ADS\,\A$.
\end{definition}

\begin{definition}[A model of ETT supports IITs]\label{sec:hasiits}
  A model of ETT supports IITs if for any signature $(sig,nat):\Sign$
  there is a
  \[
    \con_{sig}: ({sig}\,\ADS)^\A
  \]
  and an
  \[
    \elim_{sig} : (\gamma^D:({sig}\,\ADS)^\D\,\con_{sig})\ra ({sig}\,\ADS)^\S\,\con_{sig}\,\gamma^D.
  \]
\end{definition}
In other words, for any signature, we have an algebra called $\con$
(constructors) and for any displayed algebra over the constructors, we
have a section (called the eliminator).

@@

\begin{example}[An IIT supports natural numbers]
  For the signature $({nat},{natc})$ of natural numbers in \Cref{ex:signatures},
  the type of $\con_{nat}$ is

@@

  ETT, by equality reflection we also get them as definitional
  equalities.

@@

\begin{remark}
  In a model of ETT supporting the theory of ITT signatures, we can
  view elements of $\Con^\I$ as syntactic signatures. We can turn a
  $({sig},{sigc}) : \Sign$ into $\Con^\I$ by ${sig}\,\I$ and an
  $\Omega:\Con^\I$ into a $\Sign$ by
  $(\lambda M.\ll\Omega\rr_M, \lambda M\,N\,f.f\,(\ll\Omega\rr_M) =
  \ll\Omega\rr_N)$ where the equality in the second component comes
  from uniqueness of the recursor.
\end{remark}

@@

  An $a : \Tm\,\Omega\,\U$ is a sort term for a signature $\Omega$. If
  $\Omega$ is the syntactic signature for natural numbers, $a$ can
  only be $N$ ($1$ as a de Bruijn index). If $\Omega$ is the syntactic
  signature for $\Con$--$\Ty$ (\Cref{ex:signatures}), $a$ can be
  $Con$, $Ty\oldapp{empty}$,
  $Ty\oldapp({ext}\oldapp {empty}\oldapp(U\oldapp{empty}))$, and so
  on. In any case, for such an $a$, we obtain
  $(\ll a\rr_{{\A\I\C}_\Omega})^\C\,\id^\I : \Tm^\I\,\Omega\,(\El\,a)
  = \ll a\rr^\A\,(\ll\Omega\rr^\C\,\id^\I)$. That is, the algebra
  interpretation of a sort term at the constructors is equal to terms
  of that sort.

>>>>

\begin{definition}[Signature algebra, $\SignAlg$]\label{def:algebra}
  In a model of ETT, a signature algebra is an iterated $\Sigma$ type
  consisting of the following four (families of) sets, 17 operations
  and 18 equalities.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\
  & \Con && : \Set \\
  & \Ty  && : \Con\ra\Set \\
  & \Sub  && : \Con\ra\Con\ra\Set \\
  & \Tm  && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \\
  & \id && : \Sub\,\Gamma\,\Gamma \\
  & \blank\circ\blank && : \Sub\,\Theta\,\Delta\ra\Sub\,\Gamma\,\Theta\ra\Sub\,\Gamma\,\Delta \\
  & \ass && : (\sigma \circ \delta) \circ \nu = \sigma \circ (\delta \circ \nu) \\
  & \idl && : \id\circ\sigma = \sigma \\
  & \idr && : \sigma\circ\id = \sigma \\
  & \blank[\blank] && : \Ty\,\Delta\ra\Sub\,\Gamma\,\Delta\ra\Ty\,\Gamma \\
  & \blank[\blank] && : \Tm\,\Delta\,A\ra(\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma]) \\
  & [\id] && : A[\id] = A \\
  & [\circ] && : A[\sigma \circ \delta] = A[\sigma][\delta] \\
  & [\id] && : t[\id] = t \\
  & [\circ] && : t[\sigma \circ \delta] = t[\sigma][\delta] \\
  & \boldsymbol{\cdot} && : \Con \\
  & \epsilon && : \Sub\,\Gamma\,\boldsymbol{\cdot} \\
  & {\boldsymbol{\cdot}\eta} && : (\sigma : \Sub\,\Gamma\,\boldsymbol{\cdot}) \ra \sigma = \epsilon \\
  & \blank\ext\blank && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \blank,\blank && : (\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma])\ra\Sub\,\Gamma\,(\Delta\ext A) \\
  & \pi_1 && : \Sub\,\Gamma\,(\Delta\ext A)\ra \Sub\,\Gamma\,\Delta \\
  & \pi_2 && : (\sigma : \Sub\,\Gamma\,(\Delta\ext A))\ra \Tm\,\Gamma\,(A[\pi_1 \sigma]) \\
  & {\pi_1\beta} && : \pi_1(\sigma, t) = \sigma \\
  & {\pi_2\beta} && : \pi_2(\sigma, t) = t \\
  & {\pi\eta} && : (\pi_1\,\sigma, \pi_2\,\sigma) = \sigma \\
  & {,\circ} && : (\sigma, t)\circ\delta = (\sigma\circ\delta, t[\delta]) \\
  & \rlap{(2) Universe} \\
  & \U && : \Ty\,\Gamma \\
  & \El && : \Tm\,\Gamma\,\U\ra\Ty\,\Gamma \\
  & {\U[]} && : \U[\sigma] = \U \\
  & {\El[]} && : (\El\,a)[\sigma] = \El\,(a[\sigma]) \\
  & \rlap{(3) Inductive parameters} \\
  & \Pi && : (a:\Tm\,\Gamma\,\U)\ra\Ty\,(\Gamma\ext\El\,a)\ra\Ty\,\Gamma \\
  & \blank\oldapp\blank && : \Tm\,\Gamma\,(\Pi\,a\,B)\ra (u : \Tm\,\Gamma\,(\El\,a))
  \ra \Tm\,\Gamma\,(\El\,(B[\id,\,u])) \\
  & {\Pi[]} && : (\Pi\,a\,B)[\sigma] = \Pi\,(a[\sigma])\,(B[\sigma\circ\p,\q]) \\
  & {\oldapp[]} && : (t\oldapp\alpha)[\sigma] = (t[\sigma])\mathop{\oldapp}(\alpha[\sigma]) \\
  & \rlap{(4) External parameters} \\
  & \Pim && : (T:\Set)\ra(T\ra\Ty\,\Gamma)\ra\Ty\,\Gamma \\
  & \blank\appm \blank && : \Tm\,\Gamma\,(\Pim\,T\,B)\ra(\alpha:T) \ra\Tm\,\Gamma\,(B\,\alpha) \\
  & {\Pim[]} && : (\Pim\,T\,B)[\sigma] = \Pim\,T\,(\lambda \alpha.(B\,\alpha)[\sigma]) \\
  & {\appm[]} && : (t\appm\alpha)[\sigma] = (t[\sigma])\mathop{\appm}\alpha
\end{alignat*}
Given an $M:\SignAlg$, we denote its components as $\Con^M$, $\Ty^M$,
$\Sub^M$, $\Tm^M$, $\id^M$, and so on. We omit the indices if there is
only one signature algebra in scope (e.g.\ in
\Cref{def:abbrevs,ex:contexts}).
\end{definition}

\begin{definition}[Abbreviations]\label{def:abbrevs}
  For a signature algebra, we use $\wk : \Sub\,(\Gamma\ext A)\,\Gamma$
  to mean $\pi_1\,\id$. We recover de Bruijn indices by setting
  $0:= \pi_2 \,\id$ and $1+n := n[\wk]$. $\Pi\,a\,(B[\wk])$ is
  abbreviated by $a\Ra B$, $\Pim\,T\,(\lambda\_.B)$ by $T\Ram B$.
\end{definition}

\begin{example}[Example contexts in a signature algebra]\label{ex:contexts}
  Given a signature algebra, we can define a context in which
  specifies natural numbers. For readability, the an informal version
  of the same context is displayed on the right using variable names.
  \[
    \boldsymbol{\cdot}\,\ext\,\U\,\ext\,z : \El\,0\,\ext\,s : 1 \Ra \El\,1
    \hspace{4.9em}
    \boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N
  \]
  We start with the empty context $\boldsymbol{\cdot}$, then we
  declare a sort $\U$, then we declare an operator producing an
  element of the sort denoted by $\El\,0$ where $0$ is the de Bruijn
  index referring to the sort. Finally, we declare an operator which
  takes as input an element of the sort (now it became de Bruijn index
  $1$) and produces an element of the same sort. Note the assymetry of
  the function type $\Ra$: on the left there needs to be an element of
  $\U$, while on the right any type can appear (including another
  function type). This ensures strict positivity of the operators.

  Lists with elements of a given $T : \Set$ type are given by the
  following context. Here we use the function space with external
  domain $\Ram$ to include a $T$ in the signature. For readability, we
  omit the $\lambda$ and the superscripts and we don't write the
  compatibility condition. On the right we list the same signature
  with variable names.
  \begin{alignat*}{10}
    & \boldsymbol{\cdot}\,\ext\,\U\,\ext\,\El\,0\,\ext\,T\Ram 1\Ra \El\,1 \hspace{3.56em} \boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,cons : T\Ram L\Ra \El\,L
  \end{alignat*}

  The $\Con$--$\Ty$ example from \Cref{sec:intro} is given by the
  following context.
  \begin{alignat*}{5}
    & \boldsymbol{\cdot} \,\ext                                                 && \boldsymbol{\cdot} \,\ext                                                                \\
    & \U\,\ext                                                                  && Con && : \U\,\ext                                                                        \\
    & 0\Ra\U\,\ext                                                              && Ty && : Con\Ra\U\,\ext                                                                   \\
    & \El\,1\,\ext                                                              && empty && : \El\,Con\,\ext                                                                \\
    & \Pi\,2\,(2\oldapp 0 \Ra \El\,3)\ext                                       && ext && : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \Ra \El\,Con)\ext                         \\
    & \Pi\,3\,(\El\,(3\oldapp 0))\,\ext                                         && U && : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))\,\ext                                 \\
    & \Pi\,4\,(\El\,(4\oldapp(2\oldapp 0\oldapp (1\oldapp 0)))) \hspace{4.79em} && El && : \Pi\,(\Gamma : Con)\,(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp (U\oldapp\Gamma))))
  \end{alignat*}
\end{example}

The above examples are contexts in any signature algebra, and we could
take this as a definition of signature: $(M:\SignAlg)\ra\Con^M$ is the
usual Church-encoding of contexts. However (as we will see in TODO)
the notion of constructor for such signatures would be too
strong. Another approach would be to assume that there is a syntax for
signature algebras (an initial signature algebra), and then a
signature would be a context in this signature algebra. We will define
syntactic signatures using this approach in the next section
(\Cref{def:syntacticsignature}), but for now we don't want to assume
the existence of any inductive type. Instead, we will use a restricted
Church encoding which needs the notion of morphism of signatures.

The notion of morphism is determined by the notion of algebra
\cite{large_inf_qiit}, but we include it here for completeness.
\begin{definition}[Signature morphism, $\SignMor$]\label{def:morphism}
  A morphism from signature algebras $M$ to $N$ denoted
  $\SignMor\,M\,N$ consists of four functions and 17 equalities
  expressing that the functions preserve the operations of the two
  algebras. We use the same naming as in \Cref{def:algebra} and use
  superscripts to denote which algebra is meant.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\
  & \Con && : \Con^M && \ra\Con^N \\
  & \Ty  && : \Ty^M\,\Gamma && \ra\Ty^N\,(\Con\,\Gamma) \\
  & \Sub  && : \Sub^M\,\Gamma\,\Delta && \ra\Sub^N(\Con\,\Gamma)\,(\Con\,\Delta) \\
  & \Tm  && : \Tm^M\,\Gamma\,A && \ra\Tm^N\,(\Con\,\Gamma)\,(\Ty\,A) \\
  & \id && : \Sub\,\id^M && = \id^N \\
  & {\circ} && : \sigma\circ^M\delta && = \Sub\,\sigma\circ^N\Sub\,\delta \\
  & [] && : A[\sigma]^M && = \Ty\,A[\Sub\,\sigma]^N \\
  & [] && : t[\sigma]^M && = \Tm\,t[\Sub\,\sigma]^N \\
  & \boldsymbol{\cdot} && : \Con\,\boldsymbol{\cdot}^M && = \boldsymbol{\cdot}^N \\
  & \epsilon && : \Sub\,\epsilon^M && = \epsilon^N \\
  & \ext && : \Con\,(\Gamma\ext^M A) && = \Con\,\Gamma\ext^N\Ty\,A \\
  & , && : \Sub\,(\sigma,^M t) && = \Sub\,\sigma ,^N\Tm\,t \\
  & \pi_1 && : \Sub\,(\pi_1^M\,\sigma) && = \pi_1^N\,(\Sub\,\sigma) \\
  & \pi_2 && : \Tm\,(\pi_2^M\,\sigma) && = \pi_2^N\,(\Sub\,\sigma) \\
  & \rlap{(2) Universe} \\
  & \U && : \Ty\,\U^M && = \U^N \\
  & \El && : \Ty\,(\El^M\,a) && = \El^N\,(\Tm\,a) \\
  & \rlap{(3) Inductive parameters} \\
  & \Pi && : \Ty\,(\Pi^M\,a\,B) && = \Pi^N\,(\Tm\,a)\,(\Ty\,B) \\
  & \oldapp && : \Tm\,(t\mathbin{{\oldapp}^M} u) && = \Tm\,t\mathbin{{\oldapp}^N}\,\Tm\,u \\
  & \rlap{(4) External parameters} \\
  & \Pim && : \Ty\,(\Pim^M\,T\,B) && = \Pim^N\,T\,(\lambda\alpha.\Ty\,(B\,\alpha)) \\
  & \appm && : \Tm\,(t\mathbin{{\appm}^M}\alpha) && = \Tm\,t\mathbin{{\appm}^N}\alpha
\end{alignat*}
Given an $f:\SignMor\,M\,N$, we denote its first four components just
by $f_\Con$, $f_\Ty$, $f_\Sub$, $f_\Tm$ or just write $f$ if it is
clear which one is meant.
\end{definition}

We define IIT signatures using the Church encoding introduced by
Awodey, Frey and Speight \cite{DBLP:conf/lics/AwodeyFS18}. A
difference is that we avoid the usage of impredicativity as we don't
require that signatures form a signature algebra. It is also not a
problem for our use case that signatures are not in the lowest
universe.
\begin{definition}[IIT signature]
  An IIT signature is a context in an arbitrary signature algebra
  which is compatible with morphisms:
  \begin{alignat*}{6}
    & \Sign :=\,\, && \big({sig}:(M:\SignAlg)\ra\Con^M\big)\times \\
    & && \big((M\,N:\SignAlg)(f:\SignMor\,M\,N)\ra f_\Con\,({sig}\,M) = {sig}\,N \big).
  \end{alignat*}
\end{definition}
Note that as $\Sign$ refers to $\SignAlg$, it lives in a higher
universe level than $\SignAlg$. The compatibility condition says that
if we obtain an $M$-context using ${sig}$ in model $M$ and then we
transport it to $N$ using $f$, we get the same $N$-context as directly
applying ${sig}$ to $N$.

@@

\begin{example}[Example signature]\label{ex:signatures}
  Now we can formally describe the contexts given in
  \Cref{ex:contexts} as signatures. For natural numbers, we have the
  following pair of functions. The second function returns an equality
  proof which we describe using equational reasoning.
  \begin{alignat*}{5}
    & \rlap{$(nat,natc) := $} \\
    & \hspace{1em} && \big( && \rlap{$\lambda M.(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M),$} \\
    & && && \lambda M\,N\,f\,.\, && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M) = \\
    & && && && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M)\ext^N f_\Ty\,(1^N\Ra^N \El^N\,1^N) = \\
    & && && && f_\Con\,(\boldsymbol{\cdot}^M\,\ext^M\,\U^M)\,\ext^N\,f_\Ty\,(\El^M\,0^M)\ext^N f_\Tm\,1^N\Ra^M f_\Ty\,(\El^N\,1^N) = \\
    & && && && f_\Con\,\boldsymbol{\cdot}^M\,\ext^N\,f_\Ty\,\U^M\,\ext^N\,\El^N\,(f_\Tm\,0^M)\ext^N 1^M\Ra^M \El^M\,(f_\Tm\,1^N) = \\
    & && && && \boldsymbol{\cdot}^N\,\ext^N\,\U^N\,\ext^N\,\El^N\,0^N\,\ext^N\,1^N \Ra^N \El^N\,1^N \big)
  \end{alignat*}
  The first component builds the context describing natural numbers in
  $M$, the second one uses the fact that $f$ is a morphism, that is,
  it preserves all operations.

  The signatures for lists and $\Con$--$\Ty$ can be given analogously.
\end{example}

@@

\begin{definition}[The set signature algebra $\A$]\label{def:a}
  $\A : \SignAlg$ is given by the first $^\A$ components of $\ADS$
  (\Cref{def:ads}), that is, $\Con^\A := \Set$,
  $\Ty^\A\,\Gamma := \Gamma\ra\Set$,
  $\Sub^\A\,\Gamma\,\Delta := \Gamma\ra\Delta$, and so on. There is a
  morphism from $\ADS$ to $\A$ defined by $\blank^\A$ at each
  component, which we also denote by $\blank^\A : \SignMor\,\ADS\,\A$.
\end{definition}

\begin{definition}[A model of ETT supports IITs]\label{sec:hasiits}
  A model of ETT supports IITs if for any signature $(sig,sigc):\Sign$
  there is a
  \[
    \con_{sig}: ({sig}\,\ADS)^\A
  \]
  and an
  \[
    \elim_{sig} : (\gamma^D:({sig}\,\ADS)^\D\,\con_{sig})\ra ({sig}\,\ADS)^\S\,\con_{sig}\,\gamma^D.
  \]
\end{definition}
In other words, for any signature, we have an algebra called $\con$
(constructors) and for any displayed algebra over the constructors, we
have a section (called the eliminator).

@@

\begin{example}[A model of ETT supports natural numbers]
  For the signature $({nat},{natc})$ of natural numbers in
  \Cref{ex:signatures}, the type of $\con_{nat}$ is

@@

  ETT, we also get them as definitional equalities by equality
  reflection.

@@

\begin{definition}[Syntactic signatures]\label{def:syntacticsignature}
  In a model of ETT supporting the theory of ITT signatures, elements
  of $\Con^\I$ are called syntactic signatures. We can turn a
  $({sig},{sigc}) : \Sign$ into $\Con^\I$ by ${sig}\,\I$ and an
  $\Omega:\Con^\I$ into a $\Sign$ by
  $(\lambda M.\ll\Omega\rr_M, \lambda M\,N\,f.f\,(\ll\Omega\rr_M) =
  \ll\Omega\rr_N)$ where the equality in the second component comes
  from uniqueness of the recursor.
\end{definition}

@@

  An $a : \Tm\,\Omega\,\U$ is a sort term for a signature $\Omega$. If
  $\Omega$ is the syntactic signature for natural numbers, $a$ can
  only be $N$ ($1$ as a de Bruijn index). If $\Omega$ is the syntactic
  signature for $\Con$--$\Ty$ (\Cref{ex:contexts}), $a$ can be
  $Con$, $Ty\oldapp{empty}$,
  $Ty\oldapp({ext}\oldapp {empty}\oldapp(U\oldapp{empty}))$, and so
  on. In any case, for such an $a$, we obtain
  $(\ll a\rr_{{\A\I\C}_\Omega})^\C\,\id^\I : \Tm^\I\,\Omega\,(\El\,a)
  = \ll a\rr^\A\,(\ll\Omega\rr^\C\,\id^\I)$. That is, the algebra
  interpretation of a sort term at the constructors is equal to terms
  of that sort.
